Detecting Spam on Social Networking Sites

### 2012-05-08 docs/paper/paper.tex
Added table of feature statistics
update

\documentclass[preprint]{acm_proc_article-sp}
%\documentclass[preprint]{sig-alternate}
\usepackage{url}
\usepackage{graphicx,subfigure}
\usepackage{xspace}

@@

\title{Project Report: \\
Detecting Spam on Social Networking Sites}

@@

@@

\subsection{Problem}
Social networks of any significant size see constant spam, scams
and phishing attacks. The nature of these attacks can be quite diverse
and difficult detect. Marketers can spam members with unwanted
advertisements, fraudsters lure users with advance fee frauds and
other confidence tricks, while others attempt to steal user
information by directing users to external phishing pages. To make
matters more difficult, a site with global reach sees communication
among its members in a number foreign languages with varying levels
of ability. This means that much benign content shares characteristics
like misspellings, awkward phrases, etc. that might have made certain
types of common frauds and spam more easy to distinguish on US-based
(or English-language) sites.
% focus on extra impact of non-public features vs. public (bag-of-words)

@@

This paper examines in detail the types of malicious and benign
content that are encountered on social networks by analyzing
experimental data available from InterPals, an international social network
for cultural exchange and language practice. For example, the site
attracts a wide variety of financial scams, ranging from Nigerian
"419" scams to romance scams. Another prevalent problem is spam with
links to third-party websites, directing users to various porn/webcam
sites, phishing sites or various untrustworthy online marketplaces.

@@

We investigated several machine learning techniques to automatically 
detect spam and scams in the private message data, and chose to use 
train and evaluate Naive Bayes, Linear Regression, and 
Support Vector Machine (SVM) classifiers. Our implementations used 
a variety of tools, including Matlab, ScalaNLP, LIBSVM, Lucene and Spark, 
an in-memory distributed computing framework designed for machine 
learning and iterative computation.

@@

\section{Background}

@@

In contrast to research that focuses on dynamically detecting spam based on user activity, Irani et al. \cite{irani} show that static features associated with user signups on MySpace are enough to train an effective social spam classifier. They note that C4.5 decision tree algorithms provide better performance than naive Bayes in this case. As in other works, this only examines publicly available profile information collected by social honeypots. Private data collected on users including browser features, IP addresses and geographic location would conceivably improve classifier performance substantially.

@@

Many of the data mining algorithms used to detect spam and patterns of misuse on SNSs are designed with the assumption that the data and the classifier are independent. However, in the case of spam, fraud and other malicious content, users will often modify their behavior to evade detection, leading to degraded classifier performance and the need to re-train classifiers frequently. Several researchers tackle this adversarial problem. Dalvi et al. \cite{dalvi} offer a modified Naive Bayes classifier to detect and reclassify data taking into account the optimal modification strategy that an adversary could choose. Lowd and Meek \cite{lowd} provide a framework for reverse engineering a classifier to determine whether an adversary can efficiently learn enough about a classifier to effectively defeat it.

@@

At the beginning of this research project, there were just over two million messages that had been flagged by moderators as spam. We extracted the contents of exactly two million of these messages from the site's database and merged them with records of account information for the deleted users who had sent the messages. In addition to using SQL to extract much of this information, we wrote Perl scripts to clean the data, tokenize messages and prepare data for further processing. The methodology section (X) offers more details on precisely which items of data were collected and why.

@@

@@

@@

@@

\subsection{Naive Bayes}

@@

@@

>>>>

\documentclass[preprint]{acm_proc_article-sp}
%\documentclass[preprint]{sig-alternate}
\usepackage{url}
\usepackage{graphicx,subfigure}
\usepackage{xspace}
\usepackage{multirow}

@@

\title{Detecting Spam on Social Networking Sites}

@@

\maketitle
\begin{abstract}
Social networking sites (SNSs) see a variety of spam and scams targeted at 
their users. In contrast to the more limited amounts of secondary information 
available when analyzing email spam, spam on SNSs is accompanied by a wealth 
of data on the sender, which can be used to build more accurate detection mechanisms. 
We analyze public and private data from a popular social network in order to 
gain insight into the various features spam messages and the accompanying user 
accounts and profiles available to site operators. By examining the ....
\end{abstract}

@@

Social networks of any significant size see constant spam, scams
and phishing attacks. The nature of these attacks can be quite diverse
and difficult detect. Marketers can spam members with unwanted
advertisements, fraudsters lure users with advance fee frauds and
other confidence tricks, while others may attempt to steal user
information by directing users to external phishing pages. 
Sites with global reach see communication among its members in a 
number foreign languages with varying levels
of ability. This means that much benign content shares characteristics
like misspellings, awkward phrases, etc. that might have made certain
types of common frauds and spam more easy to distinguish on US-based
(or English-language) sites.
% focus on extra impact of non-public features vs. public (bag-of-words)

@@

Email spam and techniques with which to combat it has been studied 
extensively. Like
This paper examines in detail the types of malicious and benign
content that are encountered on social networks by analyzing
experimental data available from InterPals, an international social network
for cultural exchange and language practice. For example, the site
attracts a wide variety of financial scams, ranging from Nigerian
"419" scams to romance scams. Another prevalent problem is spam with
links to third-party websites, directing users to various porn/webcam
sites, phishing sites or various untrustworthy online marketplaces.

@@

We investigated several machine learning techniques to automatically 
detect spam and scams in the private message data, and chose to use 
train and evaluate Na\"ive Bayes, Linear Regression, and 
Support Vector Machine (SVM) classifiers. Our implementations used 
a variety of tools, including Matlab, ScalaNLP, LIBSVM, Lucene and Spark, 
an in-memory distributed computing framework designed for machine 
learning and iterative computation.

@@

\section{Related Work}

@@

In contrast to research that focuses on dynamically detecting spam based on user activity, Irani et al. \cite{irani} show that static features associated with user signups on MySpace are enough to train an effective social spam classifier. They note that C4.5 decision tree algorithms provide better performance than Na\"ive Bayes in this case. As in other works, this only examines publicly available profile information collected by social honeypots. Private data collected on users including browser features, IP addresses and geographic location would conceivably improve classifier performance substantially.

@@

Many of the data mining algorithms used to detect spam and patterns of misuse on SNSs are designed with the assumption that the data and the classifier are independent. However, in the case of spam, fraud and other malicious content, users will often modify their behavior to evade detection, leading to degraded classifier performance and the need to re-train classifiers frequently. Several researchers tackle this adversarial problem. Dalvi et al. \cite{dalvi} offer a modified Na\"ive Bayes classifier to detect and reclassify data taking into account the optimal modification strategy that an adversary could choose. Lowd and Meek \cite{lowd} provide a framework for reverse engineering a classifier to determine whether an adversary can efficiently learn enough about a classifier to effectively defeat it.

@@

At the beginning of this research project, slightly two million messages had been flagged by moderators as spam. We extracted the contents of exactly two million spam-labeled messages from accounts deleted by moderators between October 2011 and March 2012. As moderators can flag a user's entire list of outgoing messages as spam on deletion, the earliest sent dates of some messages in our data set begin in May 2010.

InterPals stores private messages and user account data in a number of separate MySQL tables. Account information for deleted users is stored for six months, so account details for all spam and ham accounts was available. Extracting the data of interest to this project required dumping the contents from a query with multiple joins and aggregations across 8 tables with a combined size of 154 GB. To minimize the impact on the load of the site's production database instances, we took an LVM snapshot of slave instance and cloned the database to a spare server. In addition to using SQL to extract the data, we wrote Perl scripts to clean the data, tokenize messages and prepare data for further processing. The methodology section (X) offers more details on precisely which items of data were collected and why.

@@

\textbf{Sender Photos:} We found that the number of photos that user has uploaded is also indicative of their spam or ham reputation. Spammers

@@

\begin{table}
\begin{tabular}{|l|l|r|r|r|}
    \hline
    Feature & Class & Mean & Median & Std. Dev.\\
    \hline
    \multirow{2}{*}{Unique IPs} & Spam & 11.82 & 2 & 52.52\\
    & Ham & 175.51 & 64 & 322.19\\
    \hline
    \multirow{2}{*}{Sender Age} & Spam & 33.52 & 31 & 11.09\\
    & Ham & 24.86 & 22 & 10.30\\
    \hline
    \multirow{2}{*}{Photos} & Spam & 2.09 & 1 & 8.04\\
    & Ham & 15.69 & 6 & 43.40\\
    \hline
    \multirow{2}{*}{Photo Albums} & Spam & 0.91 & 1 & 0.56\\
    & Ham & 1.89 & 1 & 2.43\\
    \hline
    \multirow{2}{*}{Friends} & Spam & 11.82 & 0 & 52.52\\
    & Ham & 23.26 & 10 & 58.91\\
    \hline
    \multirow{2}{*}{Recipient Age} & Spam & 39.98 & 41 & 12.69\\
    & Ham & 25.02 & 22 & 10.70\\
    \hline
    \multirow{2}{*}{Birth Day} & Spam & 12.31 & 10 & 8.52\\
    & Ham & 15.01 & 15 & 8.92\\
    \hline
    \multirow{2}{*}{Birth Month} & Spam & 5.85 & 5 & 3.33\\
    & Ham & 6.38 & 6 & 3.48\\
    \hline
    \multirow{2}{*}{Description Length} & Spam & 267.30 & 87 & 483.19\\
    & Ham & 439.33 & 267 & 613.26\\
    \hline
    \multirow{2}{*}{\% Fields Completed} & Spam & 22.15 & 14.00 & 30.75\\
    & Ham & 79.29 & 86.00 & 32.85\\
    \hline
\end{tabular}
\caption{Feature statistics}
\label{featsat}
\end{table}

@@

\subsection{Features for Classification}
In choosinnng

@@

Our goal in classification was not only to build an effective classifier for spam detection, but also to see how the presence of ``expert features", or features based on account and message metadata, could affect classifier accuracy. To this end, with of each of the classifiers we chose to implement, we evaluated performance on a bag-of-words representation of message data as well as a combination of bag-of-words with expert features.

%feature extraction tools: perl, spark, lucene analyzers
%- feature extraction pipeline is parallelized / distributed using Spark
%- it produces a feature matrix used later for training classifiers
%- Stage 1 generates a dictionary sorted by document frequencies in descending order
%- Stage 2 create bag-of-words representation of the examples, as well as “expert-features” along with the documents.
%
%- two feature sets: one just with bag-of-words. the other with expert features chosen by a domain expert (site owner)
%
%classification:
%- implemented 3 models: naive bayes, linear regression, and SVM.
%- used scalanlp for SVM: Pegasos: Primal Estimated sub-GrAdient SOlver for SVM Extended with Wang, Crammer, Vucetic's work for Multiclass. The optimizer runs a stochastic subgradient descent on the primal objective using batches provided.
%- used scalanlp for naive bayes
%- used matlab for linear regression with ridge regularization.
%- also attempted logistics regression with scalanlp, but training result was buggy (always output 0 probably. perhaps a bug in scalanlp)
%
%- naive bayes is super fast. accuracy around 80%. when expert features are added, accuracy dropped.
%
%- SVM is slow, and doesn’t produce very good result. Suspect we don’t have enough training data (22k training examples, 5k features)
%
%- linear regression performed the best, with an AUC of 0.xx. When expert features are added, AUC went up to 0.xx.
%

\subsection{Na\"ive Bayes}

@@

\subsection{Na\"ive Bayes}

@@

In this paper, we presented a brief overview of the breakdown of spam and scams on the InterPals web site. 
We then examined a number of statistics on both public and private data that showed that spam users, even when 
grouped together without regard to their specific angle of attack, share a number of characteristics that differentiate 
them from legitimate users. To simulate an approach 
These features allowed us to improve the accuracy of classifiers
